Carbonate chemistry at Ny-Ålesund

### 2021-04-11 awipev-CO2.Rmd
moved backwards

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
# read a file provided by Philipp for reconstructing missing pressure data
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill)
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & pressure_insitu_ctd <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & pressure_insitu_ctd <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & pressure_insitu_ctd <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)
```

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     pressure_insitu_ctd
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),       # Millero 02
   ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")

@@

# pH seaFET Final
# add despike with n=2 k = 5762 to the final ph corrected data
# despike me semble inutile, faire plutôt un simple des outliers qui, vérification faite, n'est même pas nécessaire
# Il faut par contre créér une nouvelle variable avec pH deeper than 8 m, called *_9m for consistency with salinity and temp
d_all <- d_all %>%
  #dplyr::mutate(pHint_tot_sf = despike(d_all$pHint_tot_sf, reference= "median", n=2, k=5761, replace="NA")) %>%
  dplyr::mutate(pHint_tot_sf = seafet_calib$pHint_tot_sf) %>%
  dplyr::mutate(pHint_tot_sf_1m = ifelse(pressure_insitu_ctd > 0 & pressure_insitu_ctd < 2, pHint_tot_sf, NA),
                pHint_tot_sf_3m = ifelse(pressure_insitu_ctd >= 2 & pressure_insitu_ctd < 4, pHint_tot_sf, NA),
                pHint_tot_sf_5m = ifelse(pressure_insitu_ctd >= 4 & pressure_insitu_ctd < 6, pHint_tot_sf, NA),
                pHint_tot_sf_7m = ifelse(pressure_insitu_ctd >= 6 & pressure_insitu_ctd < 8, pHint_tot_sf, NA),
                pHint_tot_sf_9m = ifelse(pressure_insitu_ctd >= 8 , pHint_tot_sf, NA),
                depth_discrete = case_when(
      !is.na(pHint_tot_sf_1m) ~ 1,
      !is.na(pHint_tot_sf_3m) ~ 3,
      !is.na(pHint_tot_sf_5m) ~ 5,
      !is.na(pHint_tot_sf_7m) ~ 7,
      !is.na(pHint_tot_sf_9m) ~ 9
      ),
      depth_discrete = as.factor(depth_discrete)
  )

# d_all <- d_all %>%
#   dplyr::mutate(ph_s_sf_temp_4 = ifelse(!is.na(pH_seafet$ph_s_sf),
#                                               pHinsi(pH = pH_seafet$ph_s_sf, 
#                          ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
#                          Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
#                          S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
#                          pHscale = "T", b="u74"), NA
#   ))

@@

```{r seafet profiles, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
# range <- d_all %>%
#   dplyr::mutate(pH_range_more_6m = ifelse(pressure_insitu_ctd > 6, pHint_tot_sf, NA)) %>%
#   group_by(montht) %>%
#   summarize(pH_range = max(pHint_tot_sf, na.rm = TRUE) - min(pHint_tot_sf, na.rm = TRUE),
#             pH_range_more_6m = max(pH_range_more_6m, na.rm = TRUE) - min(pH_range_more_6m, na.rm = TRUE),
#             pH_range_more_9m = max(pHint_tot_sf_9m, na.rm = TRUE) - min(pHint_tot_sf_9m, na.rm = TRUE)) %>%
#   ungroup()
# print(kable(range, digits = 2))

# # datetime when pH values are in but no pressure data. File sent to send to Philipp
# no_pressure <- d_all %>% 
#   dplyr::select(datetime, pressure_insitu_ctd, pHint_tot_sf) %>% 
#   dplyr::filter(is.na(pressure_insitu_ctd) & !is.na(pHint_tot_sf))
# readr::write_csv(x = no_pressure, path=paste0(path,"/fb_awipev-co2_server/ny-alesund/data//no_pressure.csv"))

@@

## Stratification
```{r stratification, echo=FALSE, message= FALSE, warning=FALSE}
depth_labels <- c("0-2", "2-4", "4-6", "6-8", "8-12")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth) %>%
  group_by(depth, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

summary(aov(tmp$sal_insitu_ctd ~ tmp$depth * tmp$montht))

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 10)
tmp_delta <- tmp %>% 
  group_by(depth, montht) %>% 
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth, values_from = sal_med) %>% 
  dplyr::rename(median0_4m = "2",
                median8_12m = "10") %>% 
  dplyr::mutate(sal_range = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity")

@@

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = density)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = c("0-4", "8-12")) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
```
Using the medians because data are missing.

@@

pangaea <- as_tibble(d_all) %>% 
  dplyr::select(datetime, pressure_insitu_ctd, sal_insitu_ctd, sal_fb, sal_mix, temp_insitu_11m, temp_insitu_ctd, temp_fb, temp_sf, temp_dur, at, ct, ph_s_sf_temp_insi, ph_s_dur_temp_fb, ph_calc_fb, pco2_fb, pco2_calc_fb, at_calc, ph_dur_corr, pHint_tot_sf) %>%
  dplyr::rename(pressure = pressure_insitu_ctd,
                s_insitu = sal_insitu_ctd,
                s_fb = sal_fb, 
                s_mix = sal_mix,
                t_11m = temp_insitu_11m,
                t_insitu = temp_insitu_ctd, 
                t_fb = temp_fb, 
                t_sf = temp_sf, 
                t_dur = temp_dur, 
                at = at, 
                at_calc =at_calc,
                ct = ct, 
                ph_s_sf_t_insi = ph_s_sf_temp_insi, 
                ph_s_dur_t_fb = ph_s_dur_temp_fb,
                pco2 = pco2_fb,
                pco2_calc = pco2_calc_fb,
                pH_sf = pHint_tot_sf,
                ph_dur = ph_dur_corr,
                ph_calc = ph_calc_fb
                )
saveRDS(file= paste0(path, "fb_data/pangaea.rds"), pangaea)
write_csv(pangaea, path = paste0(path, "fb_data/pangaea.csv"))
# Metadata Pangaea
# pressure = pressure in situ (profiler)
# s_insitu = salinity in situ (profiler)
# s_fb = salinity in FerryBox 
# s_mix = salinity in FerryBox with gaps filled by s_insitu below 8 m
# t_11m = temperature in situ at 11 m
# t_insitu = temperature in situ (profiler), 
# t_fb = temperature in FerryBox 
# t_sf = temperature SeaFET (profiler), °C
# t_dur = temperature Durafet (FerryBox)
# at = total alkalinity, discrete measurements (FerryBox)
# at_calc = total alkalinity, calculated from salinity
# ct = dissolved inorganic carbon, discrete measurements (FerryBox)
# ph_s_sf_t_insi = pH spectrophotometric SeaFET in situ (11 m)
# ph_s_dur_t_fb = pH spectrophotometric Durafet in FerryBox
# pco2 = pCO2 in Ferrybox
# pco2_calc = pCO2 calculated in the FerryBox calculated from discrete measurments of AT and CT
# pH_sf = pH SeaFET at in situ temperature (profiler)
# ph_dur = pH Durafet at FerryBox temperature (FerryBox)
# ph_calc = pH at fb temp calculated from CT and AT samples collected at the FerryBox
```

@@

# SEAFET pH
ts_pH <- d_all %>% 
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pHint_tot_sf, col = pressure_insitu_ctd), alpha=alpha, na.rm=TRUE, size=size_point) + 
  scale_colour_gradient(name = "Depth (m)", low = "#56B1F7", high = "#132B43") + # reverse colors
  guides(colour = guide_colourbar(reverse = TRUE)) + # reverse colour bar
  geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha=alpha, na.rm=TRUE, size=size_point*3) + 
  labs(title="In situ pH, spectrophotometric pH in red))",x="", y=expression(paste(pH[T]))) +
  Mytheme(size_labs = size_labs) +
  theme(legend.position = c(0.1, 0.5),
        legend.key.height = unit(0.5, 'cm'))

>>>>

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
# read a file provided by Philipp on 2021-04-09 which fills most gaps of pressure data
# depth_complete replaces all other depth and pressure variables
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill) %>% 
  dplyr::rename(depth = depth_complete)
# check_plot <- d_hour %>% 
#   ggplot() + geom_point(aes(x=datetime, y=depth, col='blue', alpha = 0.7)) +
#   geom_point(aes(x=datetime, y=depth, col='red', alpha = 0.3))
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group = ifelse( # create new variable depth
      depth <= 2,
      1,
      ifelse(
        depth > 2 & depth <= 4,
        3,
        ifelse(
          depth > 4 & depth <= 6,
          5,
          ifelse(
            depth > 6 & depth <= 8,
            7,
            ifelse(depth > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)
```

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     depth
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),       # Millero 02
   ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")

@@

# pH seaFET Final
# add despike with n=2 k = 5762 to the final ph corrected data
# despike me semble inutile, faire plutôt un simple des outliers qui, vérification faite, n'est même pas nécessaire
# Il faut par contre créér une nouvelle variable avec pH deeper than 8 m, called *_9m for consistency with salinity and temp
d_all <- d_all %>%
  #dplyr::mutate(pHint_tot_sf = despike(d_all$pHint_tot_sf, reference= "median", n=2, k=5761, replace="NA")) %>%
  dplyr::mutate(pHint_tot_sf = seafet_calib$pHint_tot_sf) %>%
  dplyr::mutate(pHint_tot_sf_1m = ifelse(depth > 0 & depth < 2, pHint_tot_sf, NA),
                pHint_tot_sf_3m = ifelse(depth >= 2 & depth < 4, pHint_tot_sf, NA),
                pHint_tot_sf_5m = ifelse(depth >= 4 & depth < 6, pHint_tot_sf, NA),
                pHint_tot_sf_7m = ifelse(depth >= 6 & depth < 8, pHint_tot_sf, NA),
                pHint_tot_sf_9m = ifelse(depth >= 8 , pHint_tot_sf, NA),
                depth_discrete = case_when(
      !is.na(pHint_tot_sf_1m) ~ 1,
      !is.na(pHint_tot_sf_3m) ~ 3,
      !is.na(pHint_tot_sf_5m) ~ 5,
      !is.na(pHint_tot_sf_7m) ~ 7,
      !is.na(pHint_tot_sf_9m) ~ 9
      ),
      depth_discrete = as.factor(depth_discrete)
  )

# d_all <- d_all %>%
#   dplyr::mutate(ph_s_sf_temp_4 = ifelse(!is.na(pH_seafet$ph_s_sf),
#                                               pHinsi(pH = pH_seafet$ph_s_sf, 
#                          ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
#                          Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
#                          S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
#                          pHscale = "T", b="u74"), NA
#   ))

@@

```{r seafet profiles, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
# range <- d_all %>%
#   dplyr::mutate(pH_range_more_6m = ifelse(depth > 6, pHint_tot_sf, NA)) %>%
#   group_by(montht) %>%
#   summarize(pH_range = max(pHint_tot_sf, na.rm = TRUE) - min(pHint_tot_sf, na.rm = TRUE),
#             pH_range_more_6m = max(pH_range_more_6m, na.rm = TRUE) - min(pH_range_more_6m, na.rm = TRUE),
#             pH_range_more_9m = max(pHint_tot_sf_9m, na.rm = TRUE) - min(pHint_tot_sf_9m, na.rm = TRUE)) %>%
#   ungroup()
# print(kable(range, digits = 2))

# # datetime when pH values are in but no pressure data. File sent to send to Philipp
# no_pressure <- d_all %>% 
#   dplyr::select(datetime, depth, pHint_tot_sf) %>% 
#   dplyr::filter(is.na(depth) & !is.na(pHint_tot_sf))
# readr::write_csv(x = no_pressure, path=paste0(path,"/fb_awipev-co2_server/ny-alesund/data//no_pressure.csv"))

@@

## Stratification
```{r stratification, echo=FALSE, message= FALSE, warning=FALSE}
depth_labels <- c("0-2", "2-4", "4-6", "6-8", "8-12")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

summary(aov(tmp$sal_insitu_ctd ~ tmp$depth_group * tmp$montht))

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth_group == 1 | depth_group == 3, 2, depth_group))) %>% 
  dplyr::filter(depth == 2 | depth == 10)
tmp_delta <- tmp %>% 
  group_by(depth, montht) %>% 
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth, values_from = sal_med) %>% 
  dplyr::rename(median0_4m = "2",
                median8_12m = "10") %>% 
  dplyr::mutate(sal_range = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity")

@@

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth), y = density)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = c("0-4", "8-12")) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
```

Using the medians because data are missing.

@@

pangaea <- as_tibble(d_all) %>% 
  dplyr::select(datetime, depth, sal_insitu_ctd, sal_fb, sal_mix, temp_insitu_11m, temp_insitu_ctd, temp_fb, temp_sf, temp_dur, at, ct, ph_s_sf_temp_insi, ph_s_dur_temp_fb, ph_calc_fb, pco2_fb, pco2_calc_fb, at_calc, ph_dur_corr, pHint_tot_sf) %>%
  dplyr::rename(pressure = depth,
                s_insitu = sal_insitu_ctd,
                s_fb = sal_fb, 
                s_mix = sal_mix,
                t_11m = temp_insitu_11m,
                t_insitu = temp_insitu_ctd, 
                t_fb = temp_fb, 
                t_sf = temp_sf, 
                t_dur = temp_dur, 
                at = at, 
                at_calc =at_calc,
                ct = ct, 
                ph_s_sf_t_insi = ph_s_sf_temp_insi, 
                ph_s_dur_t_fb = ph_s_dur_temp_fb,
                pco2 = pco2_fb,
                pco2_calc = pco2_calc_fb,
                pH_sf = pHint_tot_sf,
                ph_dur = ph_dur_corr,
                ph_calc = ph_calc_fb
                )
saveRDS(file= paste0(path, "fb_data/pangaea.rds"), pangaea)
write_csv(pangaea, path = paste0(path, "fb_data/pangaea.csv"))
# Metadata Pangaea
# pressure = pressure in situ (profiler)
# s_insitu = salinity in situ (profiler)
# s_fb = salinity in FerryBox 
# s_mix = salinity in FerryBox with gaps filled by s_insitu below 8 m
# t_11m = temperature in situ at 11 m
# t_insitu = temperature in situ (profiler), 
# t_fb = temperature in FerryBox 
# t_sf = temperature SeaFET (profiler), °C
# t_dur = temperature Durafet (FerryBox)
# at = total alkalinity, discrete measurements (FerryBox)
# at_calc = total alkalinity, calculated from salinity
# ct = dissolved inorganic carbon, discrete measurements (FerryBox)
# ph_s_sf_t_insi = pH spectrophotometric SeaFET in situ (11 m)
# ph_s_dur_t_fb = pH spectrophotometric Durafet in FerryBox
# pco2 = pCO2 in Ferrybox
# pco2_calc = pCO2 calculated in the FerryBox calculated from discrete measurments of AT and CT
# pH_sf = pH SeaFET at in situ temperature (profiler)
# ph_dur = pH Durafet at FerryBox temperature (FerryBox)
# ph_calc = pH at fb temp calculated from CT and AT samples collected at the FerryBox
```

@@

# SEAFET pH
ts_pH <- d_all %>% 
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pHint_tot_sf, col = depth), alpha=alpha, na.rm=TRUE, size=size_point) + 
  scale_colour_gradient(name = "Depth (m)", low = "#56B1F7", high = "#132B43") + # reverse colors
  guides(colour = guide_colourbar(reverse = TRUE)) + # reverse colour bar
  geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha=alpha, na.rm=TRUE, size=size_point*3) + 
  labs(title="In situ pH, spectrophotometric pH in red))",x="", y=expression(paste(pH[T]))) +
  Mytheme(size_labs = size_labs) +
  theme(legend.position = c(0.1, 0.5),
        legend.key.height = unit(0.5, 'cm'))
